---
title: "Course Project"
author: "Sofyan Abdelhay"
date: '2022-06-30'
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
```

## Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways.

## Data Source Links

The training data for this project are available here: 
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

## Load Required Packages

```{r}
## Load caret package
library(caret)
```

```{r}
## Load rpart packages
library(rpart)
library(rpart.plot)
```

```{r}
## Load rattle package
library(rattle)
```

```{r}
## Load random forest package
library(randomForest)
```

```{r}
## Load gbm package
library(gbm)
```

## Dowload Data Sets

```{r data sets}
trainURL <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testURL <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

download.file(trainURL, destfile = "training.csv")
download.file(testURL, destfile = "testing.csv")
```

## Load Data Sets

```{r}
org_training <- read.csv("training.csv")
org_testing <- read.csv("testing.csv")
dim(org_training)
```
## Convert classe predictor to factor

```{r}
org_training$classe <- as.factor(org_training$classe)
```

## Remove near-zero variance predictors

```{r}
vars <- nearZeroVar(org_training)
org_training_data <- org_training[,-vars]
org_testing_data <- org_testing[,-vars]
dim(org_testing_data)
```

## Remove predictors having more than 95% NA's

```{r}
nas <- sapply(org_training_data, function(x) mean(is.na(x))) > 0.95
org_training_data <- org_training_data[,nas == FALSE]
org_testing_data <- org_testing_data[,nas == FALSE]
dim(org_training_data)
```

## remove non-numeric predictors

```{r}
org_training_data <- org_training_data[,8:59]
org_testing_data <- org_testing_data[,8:59]
dim(org_training_data)
```

## Check training data column names

```{r}
names(org_training_data)
```

## Check testing data column names

```{r}
names(org_testing_data)
```

## Partition org_training data in training and testing sets

```{r}
inTrain <- createDataPartition(org_training_data$classe,
                               p = 0.6, list = FALSE)
training <- org_training_data[inTrain,]
testing <- org_training_data[-inTrain,]
dim(training)
```
### Create Decision Tree Model

```{r}
rpartModel <- train(classe ~ ., data = training, 
                    method = "rpart")
```

## Perform predictions on testing data set

```{r}
rpartPred <- predict(rpartModel, newdata = testing)
confusionMatrix(rpartPred, testing$classe)
```
## Plot rpart tree

```{r}
rpart.plot(rpartModel$finalModel, roundint = FALSE)
```

## Create Random Forest Model

```{r}
rfModel <- train(classe ~ ., data = training, method = "rf",
                 ntree = 25)
```

## Perform predictions on testing data set

```{r}
rfPred <- predict(rfModel, newdata = testing)
confusionMatrix(rfPred, testing$classe)
```
### Plot random forest errors

```{r}
plot(rfModel$finalModel)
```

## Create GBM Model

```{r}
gbmModel <- train(classe ~ ., data = training, method = "gbm", 
                  n.trees = 25, verbose = FALSE)
```

Model failed and did not converge

## Predict on org_testing_data usin rfModel

```{r}
predictions <- predict(rfModel, newdata = org_testing_data)
predictions
```

Random forest model achieved the best accuracy of 0.99
